---
title: "p8105_hw6_at3346"
author: "Ashley Tseng"
date: "11/27/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(measurements)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1
```{r prob1.loadtidy}
birthweight_data = 
  read_csv("./data/birthweight.csv", col_names = TRUE) %>%
  janitor::clean_names() %>% 
  mutate(
  babysex = as.factor(babysex),
  babysex = recode(
    babysex, 
    "1" = "male", 
    "2" = "female"),
  frace = as.factor(frace),
  frace = recode(
    frace, 
    "1" = "White", 
    "2" = "Black", 
    "3" = "Asian", 
    "4" = "Puerto Rican", 
    "8" = "Other", 
    "9" = "Unknown"), 
  malform = as.factor(malform),
  malform = recode(
    malform,
    "0" = "absent", 
    "1" = "present"),
  mrace = as.factor(mrace),
  mrace = recode(
    frace, 
    "1" = "White", 
    "2" = "Black", 
    "3" = "Asian", 
    "4" = "Puerto Rican", 
    "8" = "Other", 
    "9" = "Unknown"),
  delwt = round(conv_unit(delwt, "lbs", "g"), digits = 0),
  ppwt = round(conv_unit(ppwt, "lbs", "g"), digits = 0),
  wtgain = round(conv_unit(wtgain, "lbs", "g"), digits = 0),
  mheight = round(conv_unit(mheight, "inch", "cm"), digits = 0))
```


The variables I hypothesize that underly baby's birthweight include: babysex, bhead, blength, delwt, gaweeks, momage, pnumlbw, and pnumsga. I will plot each of these hypothesized variables against baby's birthweight.

```{r prob1.exploratory}
birthweight_data %>% 
  ggplot(aes(x = babysex, y = bwt)) + 
  geom_point() + 
  geom_smooth(method = 'lm')

birthweight_data %>% 
  ggplot(aes(x = bhead, y = bwt)) + 
  geom_point() + 
  geom_smooth(method = 'lm')

birthweight_data %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point() + 
  geom_smooth(method = 'lm')

birthweight_data %>% 
  ggplot(aes(x = gaweeks, y = bwt)) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```


Based on my exploratory analysis of visually inspecting the plots, I will fit a regression model with baby's sex, baby’s head circumference at birth (cm), baby’s length at birth (cm), and gestational age in weeks:
```{r prob1.reg1}
fit_og = lm(bwt ~ babysex + bhead + blength + gaweeks, data = birthweight_data)

fit_og %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 4)
```



```{r prob1.residualsplot}
birthweight_data %>% 
  modelr::add_residuals(fit_og) %>% 
  modelr::add_predictions(fit_og) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point()
```


```{r prob1.maineffects}
fit_maineff = lm(bwt ~ blength + gaweeks, data = birthweight_data)

fit_maineff %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 4)

```



```{r prob1.interactions}
fit_int = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex +  bhead * blength * babysex, data = birthweight_data)

fit_int %>%
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 4)
```



Fit models and obtain RMSE to assess prediction accuracy for different models:
```{r prob1.CV}
cv_df = 
  crossv_mc(birthweight_data, 100) 

cv_df %>% pull(train) %>% .[[1]] %>% as_tibble
cv_df %>% pull(test) %>% .[[1]] %>% as_tibble

cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(og_mod  = map(train, ~lm(bwt ~ babysex + bhead + blength + gaweeks, data = .x)),
         maineff_mod  = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         int_mod  = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))) %>% 
  mutate(rmse_og = map2_dbl(og_mod, test, ~rmse(model = .x, data = .y)),
         rmse_maineff = map2_dbl(maineff_mod, test, ~rmse(model = .x, data = .y)),
         rmse_int = map2_dbl(int_mod, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```




## Problem 2
```{r prob2.load}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r prob2.bootstrap}
set.seed(8)

boot_func = function(weather_df) {
  sample_frac(weather_df, replace = TRUE)
}

boot_samp = 
  tibble(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_func(weather_df))
  ) 
```


```{r prob2.rsquared}
bootstrap_rsquared = 
  boot_samp %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(cols = c(results)) %>% 
  select(strap_number, r.squared) %>% 
  ggplot(aes(x = r.squared)) + 
  geom_density(color = "seagreen")

bootstrap_rsquared
```



```{r prob2.logbetas}
bootstrap_logbetas = 
  boot_samp %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy)) %>% 
  unnest(cols = c(results)) %>% 
  select(strap_number, term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate) %>% 
  rename(
    "B0" = "(Intercept)",
    "B1" = "tmin") %>% 
  mutate(log_betas = log(B0*B1)) %>% 
  ggplot(aes(x = log_betas)) + 
  geom_density(color = "seagreen")

bootstrap_logbetas
```



